Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-05 app-chemistry.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

The primordial chemistry network implemented in Enzo is discussed in 
Section~\ref{sec.ov.chem} and also in
much more detail by~\citet{abel97} and~\citet{anninos97}.  
These papers describe the chemistry and cooling behavior of low-density primordial gas
($n \simeq 10^4$ and below), as well as
the steps that are necessary to obtain fast and accurate numerical solutions of the 
nonequilibrium chemical reaction rate network.  This network is extended by
~\citet{ABN02} to include the 3-body molecular hydrogen creation process, 
which becomes important at higher densities, and extends the validity of the reaction
network several more orders of magnitude in density, essentially until the gas becomes
optically thick to cooling by $H_2$ line emission. 
For the sake of completeness, and because the properties the primordial gas are so 
crucial to the results that are discussed in this work, we describe the chemistry 
network in this appendix.

Tables~\ref{table.collisional} and~\ref{table.radiative} summarize the collisional
and radiative processes solved in the Enzo nonequilibrium chemistry routines.  
\citet{abel97} show that accurate results can be obtained if several unnecessary
reactions are eliminated and the reaction network is reduced to the following:

>>>>

The primordial chemistry network implemented in \enzo\ is discussed in 
Section~\ref{sec.ov.chem} and also in
much more detail by~\citet{abel97} and~\citet{anninos97}.  
These papers describe the chemistry and cooling behavior of low-density primordial gas
($n \simeq 10^4$ and below), as well as
the steps that are necessary to obtain fast and accurate numerical solutions of the 
nonequilibrium chemical reaction rate network.  This network is extended by
~\citet{ABN02} to include the 3-body molecular hydrogen creation process, 
which becomes important at higher densities, and extends the validity of the reaction
network several more orders of magnitude in density, essentially until the gas becomes
optically thick to cooling by $H_2$ line emission. 
For the sake of completeness, and because the properties the primordial gas are so 
crucial to the results that are discussed in this work, we describe the chemistry 
network in this appendix.

Tables~\ref{table.collisional} and~\ref{table.radiative} summarize the collisional
and radiative processes solved in the \enzo\ nonequilibrium chemistry routines.  
\citet{abel97} show that accurate results can be obtained if several unnecessary
reactions are eliminated and the reaction network is reduced to the following:

### 2013-06-05 numerical-conduction.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

\enzo\ implements the equations of isotropic heat conduction in a manner
similar to that of \citet{2007ApJ...664..135P}.  
The isotropic flux of heat is given by:
\begin{equation}
\myvec{F} = -\kappa_{sp} \nabla T
\end{equation}
where $\kappa_{sp}$ is the Spitzer conduction coefficient, $4.6 \times 10^{-7}$~T$^{5/2}$ erg
s$^{-1}$~cm$^{-1}$~K$^{-1}$ \citep{1962pfig.book.....S}, here we are using a
value for the Coulomb logarithm, $\log \Lambda = 37.8$, that is
appropriate for the intracluster medium \citep{1988xrec.book.....S}.
It is quite possible that the local heat flux computed in this way 
can become unphysically large in
the high-temperature, low-density cluster regime when using this
formulation; therefore, we take into account the saturation of the heat
flux \citep{1977ApJ...211..135C} at a maximum level of
\begin{equation}
F_{sat} \simeq 0.4 n_e k_b T \left( \frac{2 k_b T}{\pi m_e} \right)^{1/2}.
\end{equation}
To ensure a smooth transition between
the Spitzer and saturated regimes, we define an effective
conductivity using the formalism of \citet{1988xrec.book.....S}
\begin{equation}
\kappa_{eff} = \frac{\kappa_{sp}}{1 + 4.2 \lambda_e / \ell_T} \; ,
\end{equation}
where $\lambda_e$ is the electron mean free path and $\ell_T \equiv T
/ |\nabla T|$ is the characteristic length scale of the local
temperature gradient.  We also assume that the conductivity of the
plasma can be described in terms of an effective conductivity, which
can be expressed as a fraction f$_{Sp}$ of the Spitzer conductivity
(where f$_{Sp} \leq 1.0$ are considered physical values).  This takes
into account physical processes below the resolution limit of the
simulation (such as tangled magnetic fields) that can suppress heat
transport.

Thermal conduction in a plasma can be strongly affected by the
presence of magnetic field lines, which may suppress heat
flow perpendicular to the magnetic field.  In that case, we allow for
heat transport only parallel to the magnetic field lines in
magnetohydrodynamic simulations.  Mathematically, this is treated as:
\begin{equation}
\myvec{F} = -\kappa_{sp} \hat{b} (\hat{b} \cdot \nabla T)
\end{equation}
where $\hat{b}$ is the unit vector denoting the direction of the
magnetic field.  As with the isotropic thermal conduction, we allow a
multiplicative factor f$_{Sp}$ to take into account the possible
behavior of magnetic fields below the resolution limit of the
simulation.

>>>>

\enzo\ implements the equations of isotropic heat conduction in a manner
similar to that of \citet{2007ApJ...664..135P}.  
The isotropic flux of heat is given by:
\begin{equation}
\myvec{F} = -\kappa_{\rm sp} \grad T
\end{equation}
where $\kappa_{\rm sp}$ is the Spitzer conduction coefficient, $4.6 \times 10^{-7}$~T$^{5/2}$ erg
s$^{-1}$~cm$^{-1}$~K$^{-1}$ \citep{1962pfig.book.....S}, here we are using a
value for the Coulomb logarithm, $\log \Lambda = 37.8$, that is
appropriate for the intracluster medium \citep{1988xrec.book.....S}.
It is quite possible that the local heat flux computed in this way 
can become unphysically large in
the high-temperature, low-density cluster regime when using this
formulation; therefore, we take into account the saturation of the heat
flux \citep{1977ApJ...211..135C} at a maximum level of
\begin{equation}
F_{sat} \simeq 0.4 n_e k_b T \left( \frac{2 k_b T}{\pi m_e} \right)^{1/2}.
\end{equation}
To ensure a smooth transition between
the Spitzer and saturated regimes, we define an effective
conductivity using the formalism of \citet{1988xrec.book.....S}
\begin{equation}
\kappa_{eff} = \frac{\kappa_{\rm sp}}{1 + 4.2 \lambda_e / \ell_T} \; ,
\end{equation}
where $\lambda_e$ is the electron mean free path and $\ell_T \equiv T
/ |\grad T|$ is the characteristic length scale of the local
temperature gradient.  We also assume that the conductivity of the
plasma can be described in terms of an effective conductivity, which
can be expressed as a fraction f$_{\rm sp}$ of the Spitzer conductivity
(where f$_{\rm sp} \leq 1.0$ are considered physical values).  This takes
into account physical processes below the resolution limit of the
simulation (such as tangled magnetic fields) that can suppress heat
transport.

Thermal conduction in a plasma can be strongly affected by the
presence of magnetic field lines, which may suppress heat
flow perpendicular to the magnetic field.  In that case, we allow for
heat transport only parallel to the magnetic field lines in
magnetohydrodynamic simulations.  Mathematically, this is treated as:
\begin{equation}
\myvec{F} = -\kappa_{\rm sp} \hat{b} (\hat{b} \cdot \grad T)
\end{equation}
where $\hat{b}$ is the unit vector denoting the direction of the
magnetic field.  As with the isotropic thermal conduction, we allow a
multiplicative factor f$_{\rm sp}$ to take into account the possible
behavior of magnetic fields below the resolution limit of the
simulation.

### 2013-06-05 numerical-mhd-ct.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

\subsection{MHD: Constrained transport}
\label{sec.num.mhd-ct}
\def\Bvec{{\bf B}}
\def\Bf{Bf}
\def\Bc{Bc}
\def\Evec{{\bf E}}
\def\Divb{\ensuremath{\nabla \cdot \Bvec}}

@@

The divergence of the magnetic field, \Divb, is
identically zero due to the fact that the evolution of the magnetic field is the
curl of a vector, and the divergence of the curl of a vector is identically
zero:
\begin{eqnarray}
\partial \Bvec/\partial t & =& \frac{1}{a} \nabla \times \Evec\\
\Divb & =& 0.
\end{eqnarray}

@@

Enzo MHD-CT has been described in great detail in \citet{Collins10},
and we direct the interested reader to that reference.  The base
Godunov method, described in \citet{Li08a}, uses spatially and
temporally second order reconstruction (both MUSCL-Hancock and
Piecewise Linear Method), a selection of Riemann solvers including
HLLC and HLLD \citep{Mignone07}.  The constrainted transport methods
are the first-order method described by \citet{Balsara99} and the
second order methods described in \citet{Gardiner05}.  The AMR
machinery is described by \citet{Balsara01} and \citet{Collins10}.

The increased complexity of the constrained transport scheme comes in
the form of area-averaged face and length-averaged edge-centered
variables, while the rest of Enzo employs predominantly
volume-averaged cell centered variables.  The magnetic field is
represented by both a face centered field, B$_f$, and a cell centered
field, B$_c$.  The electric field is edge centered.  The magnetic
field is updated in four steps: first, the Riemann problem is solved
in the traditional manner, using the cell centered field; second, an
edge-centered electric field is computed using the fluxes from the
Riemann solver; third, the curl of that electric field is used to
update the face centered field; finally, the cell centered magnetic
field is updated with an average of the face centered field.

Divergence-free AMR is somewhat more complex than the AMR employed
elsewhere in Enzo.  First, the interpolation must be constrained to be
divergence free.  Thus, all three face-centered field components are
interpolated in concert.  Second, any magnetic field information from
the previous time step must be included in the interpolation, making the
interpolation more complex than the simple parent-child relation used
for other fields.  Third, the flux correction involves more possible
grid relations than traditional AMR.  In order to circumvent this last
complexity, the electric field is projected from fine grids to parent
grids (rather than the magnetic field), and is then used to re-update
the parent magnetic field.  This is described in detail in
\citet{Balsara99} and \citet{Collins10}.

Cosmological MHD can be treated in a number of ways.  As described in Section
\red{reference}, the MUSCL with Dedner treats the cosmological dilution of
$\vecB$ with a source term, 
$$ 
\frac{\partial \vecB}{\partial t} - \frac{1}{a}  \nabla \times (\vecv \times
\vecB)  =  -\frac{\dot{a}}{2 a} \vecB. \label{eq:induction_cosmo_source}
$$
However, one can eliminate the right hand side of Equation 
\ref{eq:induction_cosmo_source} by defining
$\vecB^\prime = \vecB / \sqrt{a}.$
Then Equation \ref{eq:induction_cosmo_source} becomes
$$ 
\frac{\partial \vecB^\prime}{\partial t} - \frac{1}{a}  \nabla \times (\vecv \times
\vecB^\prime)  = 0, \label{eq:induction_cosmo_prime_nosource}
$$
and we can reuse each part of the non-cosmological algorithm (hyperbolic solver,
electric field creation, interpolation, projection, and flux correction) without
modification.

Dual energy formalism has also been incorporated in two methods, one internal
energy, similar to that described in Section \ref{sec.hydro.ppm}, and one that
uses entropy.  The additional equations are
\begin{eqnarray}
\end{eqnarray}
\begin{eqnarray}
  \frac{\partial S}{\partial t} +\frac{1}{a} \nabla \cdot (S \vecv) & = &  -
  \frac{ 3 (\gamma - 1) \dot{a}}{a} S \\
  \frac{\partial \rho e}{\partial t} + \frac{1}{a} \nabla \cdot (\rho e \vecv) &
  = & -\frac{ 3(\gamma -1 ) \dot{a}}{a} + \frac{p}{a} \nabla \cdot \vecv,
\end{eqnarray}
where $e$ is the internal energy and $S=p/\rho^{\gamma-1}$ is the entropy.  If
the ratio of thermal to kinetic energy, $\eta$, is below a certain value, the
alternative energy is used.  As in \citet{Li08a}, we use $\eta=0.008$.

Enzo's MHD-CT has been used in a variety of contexts, from galaxy
cluster evolution \citep{Xu11,Xu12}, synthetic observations of radio
halos \citep{Skillman13}, primordial star formation \citep{Xu08}, and
present-day star formation \citep{Collins11, Collins12a}.

>>>>

\subsection{MHD: Constrained transport}
\label{sec.num.mhd-ct}
\def\Bvec{{\bf B}}
\def\Bf{Bf}
\def\Bc{Bc}
\def\Evec{{\bf E}}
\def\Divb{\ensuremath{\div \Bvec}}

@@

The divergence of the magnetic field, \Divb, is
identically zero due to the fact that the evolution of the magnetic field is the
curl of a vector, and the divergence of the curl of a vector is identically
zero:
\begin{eqnarray}
\partial \Bvec/\partial t & =& \frac{1}{a} \curl \Evec\\
\Divb & =& 0.
\end{eqnarray}

@@

\enzo\ MHD-CT has been described in great detail in \citet{Collins10},
and we direct the interested reader to that reference.  The base
Godunov method, described in \citet{Li08a}, uses spatially and
temporally second order reconstruction (both MUSCL-Hancock and
Piecewise Linear Method), a selection of Riemann solvers including
HLLC and HLLD \citep{Mignone07}.  The constrainted transport methods
are the first-order method described by \citet{Balsara99} and the
second order methods described in \citet{Gardiner05}.  The AMR
machinery is described by \citet{Balsara01} and \citet{Collins10}.

The increased complexity of the constrained transport scheme comes in
the form of area-averaged face and length-averaged edge-centered
variables, while the rest of \enzo\ employs predominantly
volume-averaged cell centered variables.  The magnetic field is
represented by both a face centered field, B$_f$, and a cell centered
field, B$_c$.  The electric field is edge centered.  The magnetic
field is updated in four steps: first, the Riemann problem is solved
in the traditional manner, using the cell centered field; second, an
edge-centered electric field is computed using the fluxes from the
Riemann solver; third, the curl of that electric field is used to
update the face centered field; finally, the cell centered magnetic
field is updated with an average of the face centered field.

Divergence-free AMR is somewhat more complex than the AMR employed
elsewhere in \enzo.  First, the interpolation must be constrained to be
divergence free.  Thus, all three face-centered field components are
interpolated in concert.  Second, any magnetic field information from
the previous time step must be included in the interpolation, making the
interpolation more complex than the simple parent-child relation used
for other fields.  Third, the flux correction involves more possible
grid relations than traditional AMR.  In order to circumvent this last
complexity, the electric field is projected from fine grids to parent
grids (rather than the magnetic field), and is then used to re-update
the parent magnetic field.  This is described in detail in
\citet{Balsara99} and \citet{Collins10}.

Cosmological MHD can be treated in a number of ways.  As described in Section
\red{reference}, the MUSCL with Dedner treats the cosmological dilution of
$\vecB$ with a source term, 
$$ 
\frac{\partial \vecB}{\partial t} - \frac{1}{a}  \curl (\vecv \times
\vecB)  =  -\frac{\dot{a}}{2 a} \vecB. \label{eq:induction_cosmo_source}
$$
However, one can eliminate the right hand side of Equation 
\ref{eq:induction_cosmo_source} by defining
$\vecB^\prime = \vecB / \sqrt{a}.$
Then Equation \ref{eq:induction_cosmo_source} becomes
$$ 
\frac{\partial \vecB^\prime}{\partial t} - \frac{1}{a}  \curl (\vecv \times
\vecB^\prime)  = 0, \label{eq:induction_cosmo_prime_nosource}
$$
and we can reuse each part of the non-cosmological algorithm (hyperbolic solver,
electric field creation, interpolation, projection, and flux correction) without
modification.

Dual energy formalism has also been incorporated in two methods, one internal
energy, similar to that described in Section \ref{sec.hydro.ppm}, and one that
uses entropy.  The additional equations are
\begin{eqnarray}
\end{eqnarray}
\begin{eqnarray}
  \frac{\partial S}{\partial t} +\frac{1}{a} \div (S \vecv) & = &  -
  \frac{ 3 (\gamma - 1) \dot{a}}{a} S \\
  \frac{\partial \rho e}{\partial t} + \frac{1}{a} \div (\rho e \vecv) &
  = & -\frac{ 3(\gamma -1 ) \dot{a}}{a} + \frac{p}{a} \div \vecv,
\end{eqnarray}
where $e$ is the internal energy and $S=p/\rho^{\gamma-1}$ is the entropy.  If
the ratio of thermal to kinetic energy, $\eta$, is below a certain value, the
alternative energy is used.  As in \citet{Li08a}, we use $\eta=0.008$.

\enzo's MHD-CT has been used in a variety of contexts, from galaxy
cluster evolution \citep{Xu11,Xu12}, synthetic observations of radio
halos \citep{Skillman13}, primordial star formation \citep{Xu08}, and
present-day star formation \citep{Collins11, Collins12a}.

### 2013-06-05 numerical-mhd-dedner.tex
a little more text.
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

For the generalized Lagrange multiplier formulation of the MHD
equations \citep{}{}{2002JCoPh.175..645D}, we consider these 
conserved variables
\begin{equation}
 U = (\rho, \rho v_x, \rho v_y, \rho v_z, \rho E+B^2/2, B_x, B_y, B_z, \psi)^{T},
\end{equation} 
where $B_i$ with $i={x,y,z}$ are the three components of magnetic
fields and $\psi$ is the additional scalar field introduced in the GLM
formulation for the divergence cleaning.  The fluxes then are
\begin{eqnarray}
 F^x &=& (\rho v_x, \rho v_x^2+p+B^2/2-B_x^2, \rho v_yv_x-B_yB_x, \cr
 && \rho v_zv_x-B_zB_x, \rho ({v^2\over2} + h)v_x+B^2v_x-B_xB\cdot v, \cr
&& \psi, v_xB_y-v_yB_x, -v_zB_x+v_xB_z, c_h^2B_x)^T,\\
 F^y &=& (\rho v_y, \rho v_xv_y-B_xB_y, \rho v_y^2+p+B^2/2-B_y^2, \cr
 && \rho v_zv_y-B_zB_y, \rho ({v^2\over2} + h)v_y+B^2v_y-B_yB\cdot v, \cr
 && v_yB_z-v_zB_y,\psi,-v_xB_y+v_yB_x, c_h^2B_y)^{T}, \\
 F^z &=& (\rho v_z, \rho v_xv_z-B_xB_z, \rho v_yv_z-B_yB_z, \rho v_z^2+p+B^2/2-B_z^2, \cr
 && \rho ({v^2\over2} + h)v_z+B^2v_z-B_zB\cdot v, \cr
    &&  -v_yB_z+v_zB_y, v_zB_x-v_xB_z,\psi, c_h^2B_z)^{T},
\end{eqnarray}
where $c_h$ is a constant controlling the propagation speed and
damping rate of $\nabla\cdot B$. $h=e+p/\rho$ denotes the enthalpy.

@@

>>>>

For the generalized Lagrange multiplier formulation of the MHD
equations \citep{}{}{2002JCoPh.175..645D}, we consider these 
conserved variables
\begin{equation}
 U = (\rho, \rho v_x, \rho v_y, \rho v_z, \rho E+B^2/2, B_x, B_y, B_z, \psi)^{T},
\end{equation} 
where $B_i$ with $i={x,y,z}$ are the three components of magnetic
fields and $\psi$ is the additional scalar field introduced in the GLM
formulation for the divergence cleaning.  The fluxes then are
\begin{eqnarray}
 F^x &=& (\rho v_x, \rho v_x^2+p+B^2/2-B_x^2, \rho v_yv_x-B_yB_x, \cr
 && \rho v_zv_x-B_zB_x, \rho ({v^2\over2} + h)v_x+B^2v_x-B_xB\cdot v, \cr
&& \psi, v_xB_y-v_yB_x, -v_zB_x+v_xB_z, c_h^2B_x)^T,\\
 F^y &=& (\rho v_y, \rho v_xv_y-B_xB_y, \rho v_y^2+p+B^2/2-B_y^2, \cr
 && \rho v_zv_y-B_zB_y, \rho ({v^2\over2} + h)v_y+B^2v_y-B_yB\cdot v, \cr
 && v_yB_z-v_zB_y,\psi,-v_xB_y+v_yB_x, c_h^2B_y)^{T}, \\
 F^z &=& (\rho v_z, \rho v_xv_z-B_xB_z, \rho v_yv_z-B_yB_z, \rho v_z^2+p+B^2/2-B_z^2, \cr
 && \rho ({v^2\over2} + h)v_z+B^2v_z-B_zB\cdot v, \cr
    &&  -v_yB_z+v_zB_y, v_zB_x-v_xB_z,\psi, c_h^2B_z)^{T},
\end{eqnarray}
where $c_h$ is a constant controlling the propagation speed and
damping rate of $\div B$. $h=e+p/\rho$ denotes the enthalpy.

@@

The method is dimensionally un-split in that the fluxes are computed
for all dimensions first and the conserved quantities are updated in
one go in contrast to the Strang splitting employed in the Zeus and
PPM solvers described above.

### 2013-06-05 numerical-overview.tex
Edits to the overview section. Trying to unify notations, so several other sections are also affected.

\begin{equation}
\frac{\partial e}{\partial t} + \frac{1}{a} \vec{v}_b \cdot \nabla e = 
- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \nabla \cdot \vec{v}_b
\label{enzointenergy}
\end{equation}

@@

The \zeus\ method uses a von Neumann-Richtmeyer artificial viscosity 
to smooth shock discontinuities that may
appear in fluid flows and can cause a break-down of finite-difference
equations.  The artificial viscosity term is added in the source terms
as
\begin{eqnarray}
\rho \frac{\partial\textbf{v}}{\partial t} &=& - \nabla p - \rho \nabla \phi 
- \nabla \cdot \textbf{Q} \\
\frac{\partial e}{\partial t} &=& -p \nabla \cdot \textbf{v} - \textbf{Q} : \nabla \textbf{v}, 
\end{eqnarray}
where \textbf{v} is the baryon velocity, $\rho$ is the mass density,
$p$ is the pressure, $e$ is the internal energy density of the gas and \textbf{Q}
is the artificial viscosity stress tensor, such that:

@@

\begin{table}
\begin{center}
{\bfseries Collisional Processes}\\[1ex]
\begin{tabular}{llllllll}
(1) & H & + & e$^-$ & $\rightarrow$ & H$^+$ &+& 2e$^-$ \\
(2) & H$^+$ &+ &e$^-$ & $\rightarrow$ & H &+ &$\gamma$ \\
(3) & He &+& e$^-$ & $\rightarrow$ & He$^+$ &+& 2e$^-$  \\
(4) & He$^+$ &+& e$^-$ & $\rightarrow$ & He &+ &$\gamma$  \\
(5) & He$^{+}$ &+& e$^-$ & $\rightarrow$ & He$^{++}$ &+& 2$e^-$  \\
(6) & He$^{++}$ &+& e$^-$ & $\rightarrow$ & He$^+$ &+& $\gamma$ \\
(7) & H &+& e$^-$ &$\rightarrow$& H$^-$ &+& $\gamma$  \\
(8) & H$^-$ &+& H &$\rightarrow$ & H$_2$ & +& e$^-$ \\
(9) & H &+ &H$^+$ &$\rightarrow$ &H$_2^+$ &+ &$\gamma$ \\
(10) & H$_2^+$ &+ &H &$\rightarrow$ &$H_2$ &+ &$H^+$ \\
(11) & H$_2$ &+ &H$^+$ &$\rightarrow$ &H$_2^+$ & +& H \\
(12) & H$_2$ &+ &e$^-$ & $\rightarrow$ & 2H & + & e$^-$  \\
(13) & H$_2$ & + & H & $\rightarrow$ & 3H &   &      \\
(14) & H$^-$ & + & e$^-$ & $\rightarrow$ & H & + & 2e$^-$ \\
(15) & H$^-$ & + & H & $\rightarrow$ & 2H & + & e$^-$ \\ 
(16) & H$^-$ & + & H$^+$ & $\rightarrow$ & 2H & & \\
(17) & H$^-$ & + & H$^+$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
(18) & H$_2^+$ & + & e$^-$ & $\rightarrow$ & 2H & & \\
(19) & H$_2^+$ & + & H$^-$ & $\rightarrow$ & H$_2$ & + & H  \\
(20) & 3H & & & $\rightarrow$ & H$_2$ & + & H
\end{tabular}
\caption[]{Collisional processes solved in the Enzo nonequilibrium
primordial chemistry routines.}
\label{table.collisional}
\end{center}
\end{table}

\begin{table}
\begin{center}
{\bfseries Radiative Processes}\\[1ex]
\begin{tabular}{llllllll}
(21) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
(22) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
(23) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
(24) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
(25) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
(26) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
(27) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
(28) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
(29) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & 
\end{tabular}
\caption[]{Radiative processes solved in the Enzo nonequilibrium
primordial chemistry routines.}
\label{table.radiative}
\end{center}
\end{table}

@@

\begin{equation}
\nabla \cdot \vec{v}_b < 0,
\label{cencont}
\end{equation}

@@

\subsection{Time Stepping}\label{sec.timestepping}
\dcc{Made this its own section (not really part of the AMR, it's more
  general than that.)  Added stuff on harmonic average for 3d. See
  note below on acceleration timestep criterion being crap. Moved
  w-cycle to the previous section}
In \enzo, resolution of the equations being solved is adaptive in time 
as well as in space.  The timestep in \enzo\ is satisfied on a level-by-level
 basis by finding the largest timestep such that multiple criteria are
satisfied on each level.  The timestep criteria used by Enzo are 
(showing the one-dimensional case for clarity):

>>>>

\begin{equation}
\frac{\partial e}{\partial t} + \frac{1}{a} \vec{v}_b \cdot \grad e = 
- \frac{3(\gamma-1)\dot{a}}{a}e - \frac{p}{a \rho} \div \vec{v}_b
\label{enzointenergy}
\end{equation}

@@

The \zeus\ method uses a von Neumann-Richtmeyer artificial viscosity 
to smooth shock discontinuities that may
appear in fluid flows and can cause a break-down of finite-difference
equations.  The artificial viscosity term is added in the source terms
as
\begin{eqnarray}
\rho \frac{\partial\textbf{v}}{\partial t} &=& - \grad p - \rho \grad \phi 
- \div \textbf{Q} \\
\frac{\partial e}{\partial t} &=& -p \div \textbf{v} - \textbf{Q} : \grad \textbf{v}, 
\end{eqnarray}
where \textbf{v} is the baryon velocity, $\rho$ is the mass density,
$p$ is the pressure, $e$ is the internal energy density of the gas and \textbf{Q}
is the artificial viscosity stress tensor, such that:

@@

\begin{table}
\begin{center}
{\bfseries Collisional Processes}\\[1ex]
\begin{tabular}{llllllll}
(1) & H & + & e$^-$ & $\rightarrow$ & H$^+$ &+& 2e$^-$ \\
(2) & H$^+$ &+ &e$^-$ & $\rightarrow$ & H &+ &$\gamma$ \\
(3) & He &+& e$^-$ & $\rightarrow$ & He$^+$ &+& 2e$^-$  \\
(4) & He$^+$ &+& e$^-$ & $\rightarrow$ & He &+ &$\gamma$  \\
(5) & He$^{+}$ &+& e$^-$ & $\rightarrow$ & He$^{++}$ &+& 2$e^-$  \\
(6) & He$^{++}$ &+& e$^-$ & $\rightarrow$ & He$^+$ &+& $\gamma$ \\
(7) & H &+& e$^-$ &$\rightarrow$& H$^-$ &+& $\gamma$  \\
(8) & H$^-$ &+& H &$\rightarrow$ & H$_2$ & +& e$^-$ \\
(9) & H &+ &H$^+$ &$\rightarrow$ &H$_2^+$ &+ &$\gamma$ \\
(10) & H$_2^+$ &+ &H &$\rightarrow$ &$H_2$ &+ &$H^+$ \\
(11) & H$_2$ &+ &H$^+$ &$\rightarrow$ &H$_2^+$ & +& H \\
(12) & H$_2$ &+ &e$^-$ & $\rightarrow$ & 2H & + & e$^-$  \\
(13) & H$_2$ & + & H & $\rightarrow$ & 3H &   &      \\
(14) & H$^-$ & + & e$^-$ & $\rightarrow$ & H & + & 2e$^-$ \\
(15) & H$^-$ & + & H & $\rightarrow$ & 2H & + & e$^-$ \\ 
(16) & H$^-$ & + & H$^+$ & $\rightarrow$ & 2H & & \\
(17) & H$^-$ & + & H$^+$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
(18) & H$_2^+$ & + & e$^-$ & $\rightarrow$ & 2H & & \\
(19) & H$_2^+$ & + & H$^-$ & $\rightarrow$ & H$_2$ & + & H  \\
(20) & 3H & & & $\rightarrow$ & H$_2$ & + & H
\end{tabular}
\caption[]{Collisional processes solved in the \enzo\ nonequilibrium
primordial chemistry routines.}
\label{table.collisional}
\end{center}
\end{table}

\begin{table}
\begin{center}
{\bfseries Radiative Processes}\\[1ex]
\begin{tabular}{llllllll}
(21) & H & + & $\gamma$ & $\rightarrow$ & H$^+$ & + & e$^-$ \\
(22) & He & + & $\gamma$ & $\rightarrow$ & He$^+$ & + & e$^-$ \\
(23) & He$^+$ & + & $\gamma$ & $\rightarrow$ & He$^{++}$ & + & e$^-$ \\
(24) & H$^-$ & + & $\gamma$ & $\rightarrow$ & H & + & e$^-$ \\
(25) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^+$ & + & e$^-$ \\
(26) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & H & + & H$^+$ \\
(27) & H$_2^+$ & + & $\gamma$ & $\rightarrow$ & 2H$^+$ & + & e$^-$ \\
(28) & H$_2$ & + & $\gamma$ & $\rightarrow$ & H$_2^*$ & $\rightarrow$ & 2H \\
(29) & H$_2$ & + & $\gamma$ & $\rightarrow$ & 2H &  & 
\end{tabular}
\caption[]{Radiative processes solved in the \enzo\ nonequilibrium
primordial chemistry routines.}
\label{table.radiative}
\end{center}
\end{table}

@@

\begin{equation}
\div \vec{v}_b < 0,
\label{cencont}
\end{equation}

@@

\subsection{Time Stepping}\label{sec.timestepping}
\dcc{Made this its own section (not really part of the AMR, it's more
  general than that.)  Added stuff on harmonic average for 3d. See
  note below on acceleration timestep criterion being crap. Moved
  w-cycle to the previous section}
In \enzo, resolution of the equations being solved is adaptive in time 
as well as in space.  The timestep in \enzo\ is satisfied on a level-by-level
 basis by finding the largest timestep such that multiple criteria are
satisfied on each level.  The timestep criteria used by \enzo\ are 
(showing the one-dimensional case for clarity):
